Parallel Test

### 2025-04-16 ParallelTest.Rmd
start trying parallel processing

---
title: "MSE Scenarios for short and long lived"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

@@

## Setup for MSEs

Single area

Two surveys spring and fall

Testing alternating spring and fall

Baseline is all data all the time,

CV = empirical value from stockeff across short and long lived species

Looks like ~0.2 is reasonable for redfish in ICELAND, higher here.

Say its decent ~0.3

Eff = 100

Catch CV = 10%

Eff = 100

Project at F40% (75% F40% is HCR 1)

Assessment frequency 3 years

IID NAA

30 year spin up 30 year loop (10 assessment cycles)

F scenario for spin up? high followed by lower? Yes

Double Fmsy for 15 years and then Fmsy for 15 years

F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, changepoint = 0.5

sigma squared for NAA and recruitment?

cod assessment or redfish for these (Brian supplied redfish!)

Keep multinomial

Single fleet logistic selectivity: shift one age to the right from maturity ogive

Different selectivity by spring and fall survey: mimic some long lived species

I think this script does this

```{r}

@@

om_with_data <- update_om_fn(om, seed = 123, random = random)

@@

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
# 
# M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

@@

## Degraded data quality scenarios:

Alternate spring and fall, double CV and half Eff (50)

Fishery data, double CV (0.2) and half Eff (50)

```{r}
agg_index_sigma = input$data$agg_index_sigma
agg_index_sigma[31:60,] = 0.6 # Increase CV for both survey indices in the feedback period
index_Neff = input$data$index_Neff
index_Neff[31:60,] = 50 # Decrease ESS for both survey indices in the feedback period

#alternate years fall and spring surveys
remove_agg = TRUE # remove a aggregate index for some years
remove_agg_pointer = c(1,2) # both
remove_agg_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
remove_paa = TRUE # Also remove age comp for that index 
remove_paa_pointer = c(1,2) # both
remove_paa_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey

input <- update_input_index_info(input, agg_index_sigma, index_Neff,
                                 remove_agg, remove_agg_pointer, remove_agg_years,
                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file

agg_catch_sigma = input$data$agg_catch_sigma
agg_catch_sigma[31:60,] = 0.2 #double catch CV in the feedback period
catch_Neff = input$data$catch_Neff
catch_Neff[31:60] = 50

input <- update_input_catch_info(input, agg_catch_sigma, catch_Neff)

Generate new om and dataset, keep same assessment interval as above

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Test new EM

```{r}

mods[[2]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            # ------------------------------------------------------ #
                            # - Below is needed when making changes on data quality- #
                            # ------------------------------------------------------ #
                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
                            update_catch_info  = list(agg_catch_sigma = agg_catch_sigma, catch_Neff = catch_Neff), # Must have this!
                            # ------------------------------------------------------ #
                            # - Above is needed when making changes on data quality- #
                            # ------------------------------------------------------ #
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)

This compares the base case,good surveys in the future (1) with degraded data, alternating year seasonal surveys with lower quality sampling (catch too) in the future (2)

```{r}

par(mfrow = c(1,2))
SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB", lwd = 2) 
colors <- c("red","green","purple","orange")
lines(SSB_s1[[2]], col = colors[2],lty = 2, lwd = 2)
legend("topleft",legend = paste0("EM ", 1:2), col = c("red","green"),lty=1:2, cex=0.8)

Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,15000), main = "OM Catch", lwd = 2)
colors <- c("red","green","purple","orange")
lines(Catch_s1[[2]], col = colors[2],lty = 2, lwd = 2)

## Empirical management procedure

Index based method:  Ismooth averaging most recent 3 years of whatever surveys (rescaled to mean of each using Ismooth package)

Ismooth package is here https://github.com/cmlegault/PlanBsmooth

Do this later...

For now, do

100 replicates for each, try the parallel processing

* Base scenario 
* Alternating surveys and less age comps with 3 year assessments
* Empirical management procedures (later)

>>>>

---
title: "Parallel Test"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

@@

## Test parallel script with [Scenarios](https://noaa-edab.github.io/WHAMmse_longlived/Scenarios.html)

See Cheng's example https://lichengxue.github.io/whamMSE/05.Parallel-Computing.html

First set up the base model, all the way up to the blank mods list, but not sampling data yet

```{r base setup 3yr}
library(wham)
library(whamMSE)

@@

#om_with_data <- update_om_fn(om, seed = 123, random = random)

@@

@@

Set up folders for results

```{r}
sub.dir = "Results"
dir.create(file.path(getwd(), sub.dir), recursive = TRUE)

library(doParallel)
library(foreach)

detectCores() # check how many cores available
```

Set up clusters to run multiple sims for each scenario

```{r}
cluster <- makeCluster(5) 
registerDoParallel(cluster)

foreach (i = 1:5) %dopar% {

  om_with_data <- update_om_fn(om, seed = 123+i, random = random)

  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

  mod = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123+i,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

  saveRDS(mod, file.path(sub.dir,sprintf("Mod1_%03d.RDS",i)))

}

stopCluster(cluster)

```
Now try the same thing (base case) changing the assessment frequency

```{r}
assess.interval <- 6 #

cluster <- makeCluster(5) 
registerDoParallel(cluster)

foreach (i = 1:5) %dopar% {

  om_with_data <- update_om_fn(om, seed = 123+i, random = random)

  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

  mod = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123+i,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

  saveRDS(mod, file.path(sub.dir,sprintf("Mod2_%03d.RDS",i)))

}

stopCluster(cluster)
```
